---
title: "Juvenile Snowshoe Hares in Bonanza Creek Experimental Forest: Exploratory Report"
author: "Catherine Takata"
date: "20 Novemeber 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(lubridate)
library(effsize)
library(broom)
library(kableExtra)
library(ggbeeswarm)
```


### 1. Introduction 
A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last.


### 2. Data and analyses
A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.


#### Read in data 
```{r, include=FALSE}
bonanza_hares <- read.csv(here("data", "bonanza_hares.csv")) %>% 
  mutate(site = case_when(
    grid %in% "bonrip" ~ "Riparian",
    grid %in% "bonmat" ~ "Mature",
    grid %in% "bonbs" ~ "Lowland"
  ))
```

```{r, include=FALSE}
# in a pipe sequence, change the initial date into a different format and use mutate() function to add this new "formatted" column to the data set
# then mutate() a new column for specifically for year, in order to easily filter data 
hares_by_year <- bonanza_hares %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new))
```


### 3. Preliminary Results 

```{r, include=FALSE}
hare_counts <- hares_by_year %>% 
  count(age,year) %>% 
  filter(age == "j")

ggplot(data = hare_counts, aes(x = year, y = n)) + 
  geom_col(alpha = 0.5) +
  labs(title = "Number of Bonanza Hares per Year",
       subtitle = "Bonanza Creek (Takata, 2020)",
       x = "Year",
       y = "Number of Hares")
```


#### A. Weight based on sex comparison and descriptive statistics 
```{r}
juvenile_hares <- bonanza_hares %>% 
  filter(age == "j")


juvenile_hares_stats <- juvenile_hares %>% 
  group_by(sex) %>% 
  summarize(mean = mean(weight, na.rm = TRUE),
            sd = sd(weight, na.rm = TRUE),
            sample = n())
```

**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for recorded juvenile hare weight based on sex. Data: Kielland et al. (2017)
```{r}
juvenile_hares_stats %>% 
  kbl(col.names = c("Sex", "Mean Weight (g)", "Standard Deviation (g)", "Sample Size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```


```{r, fit.align = "center"}
ggplot(data = juvenile_hares, aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex)) +
  scale_color_manual(values = c("goldenrod2","cyan4"), na.value ="coral") +
  geom_boxplot(fill = NA, width = 0.2, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  theme_minimal() +
  labs(x = "Sex",
       y = "Weight (g)",
       title = "Female and Male Juvenile Hare Weight",
       subtitle = "Bonanza Creek, Alaksa, 1999-2012\n(Takata, 2020)",
       caption = "Figure 1") + 
  facet_wrap(~site)

```

**Figure 1.** Weight (g) observations of female and male juvenile Bonanza Snowshoe hares in three sampling areas: Lowland, Mature, and Riparian. Points correspond to indviidual hare weight observations: gold (female), cyan (male), and coral (N/A, in which sex could not be determined). Overlayed box plots endpoints indicate the variation of measurents from the 1st and 3rd quartiles; the black line and point indciate the median and mean value for each sex, respectively. Data: Data: Kielland et al. (2017)


```{r, include = FALSE}

#histogram 
ggplot(data = )
```


## 2 Sample T-test 
```{r}
# Create vectors of for female and male hares based on weight. 

weight_f <- bonanza_hares %>% 
  filter(sex == "f") %>% 
  pull(weight)

weight_m <- bonanza_hares %>% 
  filter(sex == "m") %>% 
  pull(weight)

# Inital 2 sample t test 
hare_weight_ttest <- t.test(weight_f, weight_m)
hare_weight_ttest
```


## Effect Size - Cohen's d
```{r}
cohen.d(weight_f, weight_m, na.rm = TRUE)
```

## Relationship between juvenile weight & hind foot length

```{r}
ggplot(data = juvenile_hares, aes(x = hindft, y = weight)) +
  geom_point() + 
  labs(title = "Juvenile Hares Hindfoot Length as a Function of Weight",
       subtitle = "Bonanza Creek (Takata 2020)",
       x = "Hind Foot Length (mm)",
       y = "Weight (g)")
```

## Linear Regression
```{r}
# Linear model, stored as hare_lm:
hare_lm <- lm(weight ~ hindft, data = juvenile_hares)

#return the overview: 
summary(hare_lm)
```


## Access model outputs
We can use the broom::tidy() function to get the model outputs in nice data frame format:
```{r}

```

